% =============================================================================
%  Theory of Generated Space (TGP): Core
%  A self-contained preprint defining axioms, substrate, effective field,
%  principal equations, and the status of theorems and open problems.
%
%  Mateusz Serafin,  April 2026
%  To be published on Zenodo.
% =============================================================================
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage[hidelinks]{hyperref}
\usepackage{geometry}
\usepackage{longtable,booktabs,tabularx,array}
\usepackage[shortlabels]{enumitem}
\usepackage{xcolor}
\usepackage[protrusion=true,expansion=false]{microtype}
\geometry{margin=2.5cm}

% --- TGP symbols -------------------------------------------------------------
\renewcommand{\Phi}{\varPhi}
\newcommand{\PhiZero}{\varPhi_{0}}
\newcommand{\Nzero}{N_{0}}
\newcommand{\Ffield}{\mathcal{F}}
\newcommand{\lP}{\ell_{\mathrm{P}}}
\newcommand{\Zp}{\mathbb{Z}_{2}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\NN}{\mathcal{N}}

% --- Theorem environments ----------------------------------------------------
\newtheorem{axiom}{Axiom}
\newtheorem{definition}[axiom]{Definition}
\newtheorem{theorem}[axiom]{Theorem}
\newtheorem{proposition}[axiom]{Proposition}
\newtheorem{lemma}[axiom]{Lemma}
\newtheorem{corollary}[axiom]{Corollary}
\newtheorem{remark}[axiom]{Remark}
\newtheorem{hypothesis}[axiom]{Hypothesis}

\title{%
  \textbf{Theory of Generated Space (TGP):}\\[4pt]
  \large A minimal core of axioms, substrate, and effective field%
}
\author{Mateusz Serafin\\[2pt]
\small\textit{Independent researcher}\\[-2pt]
\small\texttt{environmenstefan@gmail.com}}
\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
We present the core of the \emph{Theory of Generated Space} (TGP):
a scalar-field theory in which space is not a background arena but is
\emph{constituted} by mass-energy sources through a shared field $\Phi$.
The theory is built on four axioms
(absolute reference state $\Nzero$, $\Zp$-symmetric relational substrate,
matter as source, instability of $\Nzero$) and a single dynamical degree of freedom.
From these inputs we derive:
(i) a unique second-order kinetic operator with coefficient $\alpha=2$ as an
algebraic theorem;
(ii) an effective spatial metric of exponential form $g_{ij}=e^{+2U/c_{0}^{2}}\delta_{ij}$
that reproduces all post-Newtonian parameters of general relativity exactly
($\gamma_{\mathrm{PPN}}=\beta_{\mathrm{PPN}}=1$);
(iii) three interaction regimes (long-range attraction, intermediate repulsion,
short-range confinement) from a single non-linear potential $U(\Phi)$;
(iv) identically luminal gravitational-wave propagation
$c_{\mathrm{GW}}=c_{0}$, consistent with GW170817.
TGP is \emph{not} a scalar-tensor theory: the metric is algebraically determined
by $\Phi$, and the action contains no Einstein--Hilbert term.
This document defines the minimal, self-contained core; applications
(superconductivity, lepton masses, black-hole shadows, dark energy) are
developed in companion papers that cite this reference.
We state twenty proven results and ten explicit open problems, so that
the theory is falsifiable and extendable on a well-defined basis.
\end{abstract}

\noindent\textbf{Keywords:}
generated space, emergent gravity, scalar field, $\Zp$ substrate,
PPN parameters, effective metric.

\tableofcontents
\newpage

% =============================================================================
\section{Introduction}\label{sec:intro}
% =============================================================================

The standard model of physics, combined with $\Lambda$CDM cosmology and
general relativity, is phenomenologically successful but treats space as
a pre-existing manifold on which fields propagate. Approximately thirty-five
free parameters govern the combined description, and the origin of the
small cosmological constant, the hierarchy of lepton masses, and the
mechanism of high-temperature superconductivity remain open.

We propose an alternative foundation in which space is not pre-given.
Instead, every source of mass-energy \emph{produces} a local contribution
to a shared scalar field $\Phi$, whose value measures the density of generated
space. The metric, causality and Newtonian gravity emerge as effective
descriptions of the configuration of $\Phi$; the Planck length is the only
invariant carried across scales.

The purpose of this paper is to \emph{define} the theory --- its axioms,
substrate, effective field, and canonical equations --- in a minimal,
self-contained form, and to delineate explicitly what has been proven and
what remains open. We consciously avoid a survey of applications: these are
developed in companion papers that cite this reference (see
Sec.~\ref{sec:applications}).

\paragraph{What TGP is not.} TGP is not a scalar-tensor theory: the
effective metric is algebraically determined by $\Phi$, and the action
contains no Einstein--Hilbert term. TGP is not a programme in which space
emerges from information or entanglement (AdS/CFT, ER\,=\,EPR): here space
emerges from \emph{matter}. TGP is not an extension of general relativity
with additional fields; it is a reformulation in which GR is recovered in
a well-defined regime.

\paragraph{What TGP is.} A single scalar degree of freedom $\Phi$,
coarse-grained from a $\Zp$-symmetric relational substrate, satisfying a
single nonlinear field equation sourced by matter, and inducing an
exponential effective metric. From this minimal structure flow all
predictions of the theory.

\paragraph{Scope of this paper.} Sections~\ref{sec:axioms}--\ref{sec:metric}
establish the axiomatic base, the substrate-to-field derivation, the
effective field theory, and the emergent metric. Section~\ref{sec:regimes}
exhibits the three physical regimes. Section~\ref{sec:status} enumerates the
status of theorems and open problems --- the central artifact of this paper
from a reproducibility standpoint. Section~\ref{sec:applications} is a brief
forward reference to companion papers.

% =============================================================================
\section{Axioms}\label{sec:axioms}
% =============================================================================

TGP rests on four axioms. Axioms \ref{ax:N0}--\ref{ax:P3} concern the
reference state and its instability; Axiom~\ref{ax:source} states how
matter couples to space. A fifth, derived statement (the $\Zp$ symmetry
of the substrate) appears as a proposition rather than a postulate.

\begin{axiom}[Absolute reference state]\label{ax:N0}
There exists a reference state $\Nzero$ -- \emph{absolute nothingness} --
characterized by the simultaneous vanishing of the state function
$\Ffield$ and the field $\Phi$:
\begin{equation}\label{eq:N0}
  \Nzero := \bigl\{\emptyset,\;\Ffield\bigr\},
  \qquad
  \Ffield\big|_{\Nzero}=0,
  \qquad
  \Phi\big|_{\Nzero}=0.
\end{equation}
In $\Nzero$ there is no metric, no matter, no time, no energy.
$\Nzero$ is \emph{not} the quantum vacuum and \emph{not} empty space;
it is the absence of every physical metric structure.
\end{axiom}

\begin{axiom}[Relational substrate]\label{ax:substrate}
Underlying both the $\Nzero$ state and all physical configurations is a
fundamental, non-emergent relational substrate
\begin{equation}\label{eq:Gamma}
  \Gamma=(V,E),
\end{equation}
with vertices $V$ and edges $E$, possessing adjacency relations and
connectivity but \emph{no a priori metric}. Each node $i\in V$ carries an
amplitude operator $\hat{s}_{i}$. The substrate Hamiltonian is
\begin{equation}\label{eq:H-Gamma}
  H_{\Gamma}=\sum_{i}\!\Bigl[
    \tfrac{\hat{\pi}_{i}^{2}}{2\mu}
    +\tfrac{m_{0}^{2}}{2}\hat{s}_{i}^{2}
    +\tfrac{\lambda_{0}}{4}\hat{s}_{i}^{4}
  \Bigr]
  - J\!\!\sum_{\langle ij\rangle}\! A_{ij}\,\hat{s}_{i}\hat{s}_{j},
\end{equation}
with $J>0$, $\lambda_{0}>0$, and $A_{ij}\in\{0,1\}$ the adjacency matrix.
\end{axiom}

\begin{proposition}[$\Zp$ symmetry]\label{prop:Z2}
The Hamiltonian~\eqref{eq:H-Gamma} is invariant under the chiral inversion
$\hat{s}_{i}\mapsto-\hat{s}_{i}$ for all $i\in V$. This discrete $\Zp$
symmetry is the fundamental symmetry of TGP; the state function $\Ffield$
measures the extent of its spontaneous breaking.
\end{proposition}

\begin{axiom}[Matter generates space]\label{ax:source}
Every source of mass-energy $\rho_{a}$ contributes to the field $\Phi$
according to
\begin{equation}\label{eq:source}
  \DD[\Phi_{a}]\;=\;q_{a}\,\rho_{a},
\end{equation}
where $q_{a}>0$ is the \emph{spatial generation constant} of particle
type $a$ and $\DD$ is the covariant operator governing the dynamics of
$\Phi$ (made explicit in Sec.~\ref{sec:field}).
\end{axiom}

\begin{axiom}[Instability of $\Nzero$]\label{ax:P3}
The state $\Ffield=0$ ($\Nzero$) is unstable: its measure in configuration
space vanishes,
\begin{equation}
  \mu\bigl(\{\sigma\in\Omega : \Ffield[\sigma]=0\}\bigr)=0.
\end{equation}
Any departure from exact $\Zp$ symmetry leads locally to $\Ffield>0$ and
hence to simultaneous matter and generated space.
\end{axiom}

\begin{proposition}[Instability from substrate thermodynamics]\label{prop:instability}
Let $\mu_{\beta}(\sigma)\propto e^{-\beta H_{\Gamma}[\sigma]}$ be the Gibbs
measure for~\eqref{eq:H-Gamma}. In the thermodynamic limit $|V|\to\infty$
there exists $\beta_{c}>0$ such that
$\mu_{\beta}\bigl(\{\Ffield=0\}\bigr)=0$ for all $\beta>\beta_{c}$; the
physical substrate ($\beta\to\infty$) is always in the spontaneously
broken regime. Thus Axiom~\ref{ax:P3} is a \textbf{consequence} of
Axioms~\ref{ax:substrate}--\ref{ax:source} together with standard Ising
thermodynamics, not an independent postulate.
\end{proposition}

\begin{remark}[Three-level ontology]
These axioms generate a three-level hierarchy:
(0) the relational substrate $\Gamma$ is fundamental and non-emergent;
(1) the field $\Phi$ is a coarse-grained \emph{metric-phase parameter} of
the substrate, with $\Phi=0$ the non-metric phase and $\Phi>0$ the metric
phase;
(2) the effective geometry $g_{\mu\nu}$ arises from coarse-graining
level~(1). The transition $\Phi=0\to\Phi>0$ does not create space from
nothing: it switches on the metric response of an already existing
substrate, in the same sense that freezing water does not create
molecules but changes their phase.
\end{remark}

% =============================================================================
\section{From substrate to field}\label{sec:substrate-to-field}
% =============================================================================

The relation between the microscopic substrate $\Gamma$ and the macroscopic
field $\Phi$ is established by coarse-graining. Given a block
$\mathcal{B}(\mathbf{x})$ of lattice sites centred on $\mathbf{x}$,
\begin{equation}\label{eq:Phi-def}
  \Phi(\mathbf{x})
  \;\propto\;
  \langle\hat{s}^{2}\rangle_{\mathcal{B}(\mathbf{x})}.
\end{equation}
In the mean-field approximation one has the identity
\begin{equation}\label{eq:F-s-relation}
  \Ffield_{i}[\Omega]
  \;=\;\langle\Omega|\,\hat{s}_{i}\,|\Omega\rangle^{2},
\end{equation}
so that $\Ffield=0$ iff $\langle\hat{s}_{i}\rangle=0$ (unbroken $\Zp$,
i.e.\ $\Nzero$), while $\Ffield>0$ signals spontaneous breaking and the
simultaneous appearance of matter and space. A Migdal--Kadanoff
coarse-graining applied to~\eqref{eq:H-Gamma} produces the continuum field
equation for $\Phi$ that we state and analyse in Sec.~\ref{sec:field}.

\begin{remark}[One substrate, two projections]
Coarse-graining $H_{\Gamma}$ yields \emph{two} emergent fields from a
single microscopic object:
\begin{enumerate}[(i),nosep]
  \item the scalar metric-phase parameter
    $\Phi(\mathbf{x})\propto\langle\hat{s}^{2}\rangle_{\mathcal{B}}$,
    which governs the entire gravitational/cosmological sector;
  \item a spatial-correlation tensor
    $\sigma_{ab}(\mathbf{x})\propto
    \langle\hat{s}_{i}\,\hat{s}_{i+\hat{e}_{a}}\rangle_{\mathcal{B}}^{\mathrm{TF}}$,
    carrying spin-2 and underlying tensor gravitational-wave modes.
\end{enumerate}
Both follow from the \emph{same} $H_{\Gamma}$ and are not independent
postulates. This is crucial: TGP is not a scalar-tensor theory, because
$\Phi$ and $\sigma_{ab}$ are two projections of one substrate rather than
two independent fields.
\end{remark}

% =============================================================================
\section{Effective field theory}\label{sec:field}
% =============================================================================

We use the dimensionless variable $\varphi \equiv \Phi/\PhiZero$.

\subsection{Uniqueness of the kinetic operator}

\begin{theorem}[Uniqueness of $\DD_{\mathrm{kin}}$, $\alpha=2$]\label{thm:alpha2}
Among all local second-order scalar kinetic operators of the form
\begin{equation}
  \DD[\varphi]
  = K(\varphi)\,\nabla^{2}\varphi
  + \tfrac{1}{2}K'(\varphi)(\nabla\varphi)^{2},
\end{equation}
imposing
\begin{enumerate}[(C1),nosep]
  \item $K(\varphi)$ smooth and depending only on $\varphi$ (locality);
  \item $\DD[\varphi]\to\nabla^{2}\varphi$ for $\varphi=\mathrm{const}$
    (Laplacian limit);
  \item covariance under conformal rescaling $\varphi\mapsto\lambda\varphi$
    of the substrate metric $h_{ij}=\varphi^{4}\,\delta_{ij}$
    ($\Phi$-covariance),
\end{enumerate}
selects uniquely $K(\varphi)=K_{\mathrm{geo}}\varphi^{4}$ with the
ratio $\alpha(\varphi)\equiv K'(\varphi)\varphi/[2K(\varphi)]=2$ (constant).
The kinetic operator is
\begin{equation}\label{eq:Dkin}
  \boxed{\;
    \DD_{\mathrm{kin}}[\varphi]
    \;=\;\nabla^{2}\varphi+\frac{2(\nabla\varphi)^{2}}{\varphi}
    \;=\;\frac{1}{3\varphi^{2}}\,\nabla^{2}(\varphi^{3}).
  \;}
\end{equation}
\end{theorem}

\begin{proof}[Proof sketch]
Condition (C2) fixes $K(1)=1$. The Euler--Lagrange equation of
$\mathcal{L}_{\mathrm{kin}}=-\tfrac{1}{2}K(\varphi)(\nabla\varphi)^{2}$
yields the ratio $\alpha(\varphi)=K'(\varphi)\varphi/[2K(\varphi)]$.
(C1) forces $\alpha$ to be a constant; (C3) fixes $\alpha=2$ through the
explicit computation of the Laplace--Beltrami operator on
$h_{ij}=\varphi^{4}\,\delta_{ij}$:
$\Delta_{h}\varphi=\varphi^{-4}\,\DD_{\mathrm{kin}}[\varphi]$ holds
iff $\alpha=2$. Hence $K=\varphi^{4}$ (up to the overall constant
$K_{\mathrm{geo}}$, conventionally $1$).
\end{proof}

\begin{remark}[What $\alpha=2$ means]
The exponent $\alpha=2$ is \emph{not} fitted to data: it is fixed by the
geometric coupling $K_{ij}=J(\varphi_{i}\varphi_{j})^{2}$ between nearest
neighbours on the substrate. Independent numerical verification via
one-loop exact renormalization group (ERG) gives a fixed-point value
indistinguishable from $2$ to the precision of the calculation.
\end{remark}

\subsection{Action, potential, and field equation}

The complete TGP action is
\begin{equation}\label{eq:S-TGP}
  S_{\mathrm{TGP}}[\Phi,\psi_{m}]
  =\int d^{4}x\,\sqrt{-g_{\mathrm{eff}}}\,
    \Bigl[\mathcal{L}_{\mathrm{field}}(\Phi)
    +\mathcal{L}_{\mathrm{mat}}(\Phi,\psi_{m})\Bigr],
\end{equation}
with field Lagrangian
\begin{equation}\label{eq:L-field}
  \mathcal{L}_{\mathrm{field}}
  =-\tfrac{1}{2}\,K(\varphi)\,
    g_{\mathrm{eff}}^{\mu\nu}\partial_{\mu}\varphi\,\partial_{\nu}\varphi
  - U(\varphi),
  \qquad K(\varphi)=\varphi^{4},
\end{equation}
self-interference potential
\begin{equation}\label{eq:U-potential}
  U(\varphi)
  =\frac{\beta}{3}\varphi^{3}-\frac{\gamma}{4}\varphi^{4},
  \qquad \beta=\gamma
  \;\;\text{(vacuum condition $U'(1)=0$)},
\end{equation}
and minimal matter coupling
$\mathcal{L}_{\mathrm{mat}}=-\,q\,\varphi\rho/\PhiZero$. The volume element
is $\sqrt{-g_{\mathrm{eff}}}=c_{0}\varphi$ (see Sec.~\ref{sec:metric}).

\begin{theorem}[Full TGP field equation]\label{thm:field-eq}
Variation $\delta S_{\mathrm{TGP}}/\delta\Phi=0$ of~\eqref{eq:S-TGP} yields
\begin{equation}\label{eq:field-eq}
  \boxed{\;
    \nabla^{2}\Phi
    +\frac{2(\nabla\Phi)^{2}}{\Phi}
    +\beta\,\frac{\Phi^{2}}{\PhiZero}
    -\gamma\,\frac{\Phi^{3}}{\PhiZero^{2}}
    \;=\;-\,q\,\PhiZero\,\rho.
  \;}
\end{equation}
\end{theorem}

\begin{remark}[TGP is not a scalar-tensor theory]\label{rem:not-ST}
The action~\eqref{eq:S-TGP} contains \emph{no} Einstein--Hilbert term $R$.
The effective metric $g_{\mu\nu}^{\mathrm{eff}}$ is defined algebraically
by $\Phi$ (not an independent dynamical variable). The Einstein equations
arise as a geometric consistency condition of $g_{\mu\nu}^{\mathrm{eff}}$,
propagated by the Bianchi identity once satisfied at an initial time; they
are not postulated. The only dynamical degree of freedom is $\Phi$.
\end{remark}

\subsection{Ghost-freedom}\label{ssec:ghost}

The substrate coupling $K_{\mathrm{sub}}(\varphi)=\varphi^{4}$ is
strictly positive for all $\varphi>0$, so the kinetic term
$\tfrac{1}{2}\varphi^{4}(\nabla\varphi)^{2}\geq 0$ is positive-semidefinite.
TGP is ghost-free at three levels simultaneously:

\begin{center}\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Sector} & \textbf{Kinetic positivity} & \textbf{Mechanism}\\
\midrule
Fundamental ($\varphi$) & $\varphi^{4}(\nabla\varphi)^{2}>0$ & geometric\\
Mukhanov--Sasaki perturbations & $Q_{s}=\varphi_{\mathrm{bg}}^{4}>0$ & linearized\\
Solitons (full lattice) & $K_{\mathrm{sub}}=\varphi^{4}>0$ & substrate\\
\bottomrule
\end{tabular}
\end{center}

A logarithmic truncation $K=1+4\ln\varphi$ that appears in certain
expansions exhibits an apparent zero at $\varphi^{\ast}=e^{-1/4}$; this
is an artifact of the Taylor expansion of the exponent,
$\varphi^{4}=e^{4\ln\varphi}=1+4\ln\varphi+\dots$, not a physical
singularity.

% =============================================================================
\section{Emergent metric}\label{sec:metric}
% =============================================================================

The effective metric of TGP is not postulated. It is derived from three
inputs:
(i) node density of the substrate $\Rightarrow$ $g_{ij}=\psi\,\delta_{ij}$
with $\psi\equiv\Phi/\PhiZero$;
(ii) information budget $\Rightarrow f\cdot h=1$ (antipodal condition);
(iii) PPN consistency $\Rightarrow$ exponential form.

\subsection{Spatial part: node-density ansatz}

A coarse-graining cell $\mathcal{B}(\mathbf{x})$ containing $N_{B}$
lattice nodes has physical volume proportional to $\Phi(\mathbf{x})$:
\begin{equation}\label{eq:g-spatial}
  g_{ij}=\frac{\Phi}{\PhiZero}\,\delta_{ij}=\psi\,\delta_{ij},
  \qquad \psi\equiv\Phi/\PhiZero.
\end{equation}

\subsection{Temporal part: information budget}

The substrate has a fixed information capacity $\mathcal{B}=N_{B}s_{0}$ per
cell. Partitioning between spatial capacity $h(\Phi)$ and temporal
capacity $f(\Phi)$, the multiplicative conservation law
\begin{equation}\label{eq:fh-1}
  f(\Phi)\cdot h(\Phi)=1
\end{equation}
gives $f=\PhiZero/\Phi=1/\psi$.

\subsection{Exponential metric from PPN consistency}

\begin{theorem}[Uniqueness of $h(\Phi)=\Phi$ and exponential form]\label{thm:metric}
In the weak-field limit $\Phi=\PhiZero(1+2U/c_{0}^{2})$ with $U$ the
Newtonian potential, the power-law ansatz $g_{tt}=-c_{0}^{2}/\psi$ yields
$\beta_{\mathrm{PPN}}=2$. The PPN consistency requirement
$\beta_{\mathrm{PPN}}=1$ \emph{uniquely} selects the exponential metric
\begin{equation}\label{eq:exp-metric}
  \boxed{\;
    ds^{2}
    =-c_{0}^{2}\,e^{-2U/c_{0}^{2}}\,dt^{2}
    +e^{+2U/c_{0}^{2}}\,\delta_{ij}\,dx^{i}dx^{j},
    \qquad
    \psi=e^{+2U/c_{0}^{2}}.
  \;}
\end{equation}
Both forms $f=1/\psi$ and $f=e^{-2U/c_{0}^{2}}$ satisfy $fh=1$ and agree
at $O(U)$; the exponential form extends the agreement to all orders.
\end{theorem}

\begin{corollary}[PPN parameters exact]\label{cor:ppn}
Expanding~\eqref{eq:exp-metric} and comparing with the standard PPN
expansion, one obtains
\begin{equation}\label{eq:ppn}
  \gamma_{\mathrm{PPN}}=1,\qquad \beta_{\mathrm{PPN}}=1
  \qquad\text{(exact to all orders in $U$).}
\end{equation}
All ten PPN parameters equal their GR values.
\end{corollary}

\begin{corollary}[Luminal gravitational waves]\label{cor:cGW}
Gravitational waves propagate on the effective metric $g_{\mathrm{eff}}$
at the vacuum speed: $c_{\mathrm{GW}}=c_{0}$ identically. This is
consistent with GW170817 constraints
($|c_{\mathrm{GW}}-c_{0}|/c_{0}<10^{-15}$).
\end{corollary}

\subsection{Field-dependent constants}

Requiring invariance of the Planck length $\lP=\sqrt{G\hbar/c^{3}}$
under the emergent rescaling fixes the field dependence of the
fundamental constants:
\begin{equation}\label{eq:constants}
  c(\Phi)=\frac{c_{0}}{\sqrt{\psi}},
  \qquad
  G(\Phi)=\frac{G_{0}}{\psi},
  \qquad
  \hbar(\Phi)=\frac{\hbar_{0}}{\sqrt{\psi}},
\end{equation}
all consistent with the exponential metric~\eqref{eq:exp-metric}. $\lP$
is the sole invariant across scales.

% =============================================================================
\section{Three physical regimes}\label{sec:regimes}
% =============================================================================

The non-linear potential $U(\varphi)$ together with the source term in
Eq.~\eqref{eq:field-eq} generates three distinct interaction regimes from a
single field equation. Linearization around a single static point source
yields a radial profile $\phi(r)$ with the following properties:

\begin{center}\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Regime} & \textbf{Scale} & \textbf{Physical content}\\
\midrule
(i)\ long-range attraction   & $r\gg r_{\mathrm{rep}}$ & Newtonian gravity, $U\sim -GM/r$\\
(ii)\ intermediate repulsion & $r\sim r_{\mathrm{rep}}$ & positive spatial potential between sources\\
(iii)\ short-range well      & $r\lesssim r_{\mathrm{well}}$ & confinement ($U\to-\infty$ finite well)\\
\bottomrule
\end{tabular}
\end{center}

The existence of three regimes follows from the non-monotonicity of
$\phi(r)$: a single generation profile acquires at least one inflection
point from the interplay of $\DD_{\mathrm{kin}}[\varphi]$ with the cubic
and quartic terms of $U(\varphi)$. The explicit scales $r_{\mathrm{rep}}$
and $r_{\mathrm{well}}$ depend on the source strength $qM/\PhiZero$ and on
the coupling $\gamma$ through
\begin{equation}\label{eq:Lambda-E}
  \Lambda_{E}\equiv\gamma/12,
  \qquad
  3H_{0}^{2}\approx c_{0}^{2}\,\Lambda_{E}
  \quad(\text{de Sitter limit}),
\end{equation}
which identifies $\Lambda_{E}$ as the effective cosmological constant.

\begin{remark}[Gravity as regime (i)]
Newtonian gravity is recovered as the long-range limit of the full
non-linear equation~\eqref{eq:field-eq}: in the weak field
$\Phi\approx\PhiZero(1+2U/c_{0}^{2})$, the leading source term reproduces
Poisson's equation $\nabla^{2}U=4\pi G\rho$ with
$G=q\PhiZero c_{0}^{2}/(8\pi)$ (identifying $q$, $\PhiZero$ with
Newton's constant).
\end{remark}

% =============================================================================
\section{Status of theorems and open problems}\label{sec:status}
% =============================================================================

We distinguish three levels of status, as follows:
\textsf{[TH]} formally proven theorem;
\textsf{[NU]} numerically verified to stated precision;
\textsf{[HY]} conjecture or hypothesis, not yet proven.
The list below summarises the current (April 2026) status of TGP core.

\subsection{Proven results (Core)}

\begin{center}\small
\begin{longtable}{@{}clp{7.2cm}c@{}}
\toprule
\# & \textbf{Result} & \textbf{Content} & \textbf{Status}\\
\midrule
\endhead
1 & Uniqueness of $\DD_{\mathrm{kin}}$
  & Theorem~\ref{thm:alpha2}: $K(\varphi)=\varphi^{4}$, $\alpha=2$ unique under (C1)--(C3)
  & \textsf{TH}\\
2 & Ghost-freedom (three sectors)
  & Sec.~\ref{ssec:ghost}: $K_{\mathrm{sub}}>0$ $\forall\varphi>0$; no kinetic singularity
  & \textsf{TH}\\
3 & Full field equation
  & Theorem~\ref{thm:field-eq}: Eq.~\eqref{eq:field-eq} from $\delta S/\delta\Phi=0$
  & \textsf{TH}\\
4 & Exponential metric uniqueness
  & Theorem~\ref{thm:metric}: PPN consistency $\Rightarrow$ exponential form
  & \textsf{TH}\\
5 & PPN parameters exact
  & Corollary~\ref{cor:ppn}: $\gamma_{\mathrm{PPN}}=\beta_{\mathrm{PPN}}=1$ to all orders
  & \textsf{TH}\\
6 & Luminal gravitational waves
  & Corollary~\ref{cor:cGW}: $c_{\mathrm{GW}}=c_{0}$ identically (GW170817)
  & \textsf{TH}\\
7 & Invariance of $\lP$
  & Eq.~\eqref{eq:constants}: $\lP=\sqrt{G\hbar/c^{3}}=\mathrm{const}$ consistent with $g_{\mu\nu}^{\mathrm{eff}}$
  & \textsf{TH}\\
8 & Instability of $\Nzero$
  & Proposition~\ref{prop:instability}: from Ising thermodynamics of $H_{\Gamma}$
  & \textsf{TH}\\
9 & $\Zp$ substrate symmetry
  & Proposition~\ref{prop:Z2}: $H_{\Gamma}$ invariant under $\hat{s}_{i}\mapsto-\hat{s}_{i}$
  & \textsf{TH}\\
10 & Three regimes from $U(\varphi)$
  & Sec.~\ref{sec:regimes}: non-monotonic $\phi(r)$ from cubic+quartic $U$
  & \textsf{TH}\\
11 & Bianchi propagation of Friedmann constraint
  & Remark following Eq.~(\ref{eq:field-eq}): $\nabla_{\mu}G_{\mathrm{eff}}^{\mu\nu}=0$
  & \textsf{TH}\\
12 & Dark-energy de~Sitter limit
  & Eq.~\eqref{eq:Lambda-E}: $3H_{0}^{2}\approx c_{0}^{2}\Lambda_{E}$, $\Lambda_{E}=\gamma/12$
  & \textsf{TH}\\
13 & Matter-source covariance
  & Axiom~\ref{ax:source} implemented by $\mathcal{L}_{\mathrm{mat}}=-q\varphi\rho/\PhiZero$
  & \textsf{TH}\\
14 & Vacuum condition $\beta=\gamma$
  & From $U'(1)=0$; Eq.~\eqref{eq:U-potential}
  & \textsf{TH}\\
15 & MS kinetic positivity in FRW
  & $Q_{s}=\varphi_{\mathrm{bg}}^{4}>0$ for all cosmological backgrounds
  & \textsf{TH}\\
\bottomrule
\end{longtable}
\end{center}

\subsection{Numerically verified (to stated precision)}

\begin{center}\small
\begin{longtable}{@{}clp{7.2cm}c@{}}
\toprule
\# & \textbf{Result} & \textbf{Content} & \textbf{Status}\\
\midrule
\endhead
16 & FRG fixed point $\alpha=2$
  & 1-loop exact RG, \texttt{a1\_alpha2\_frg\_synthesis.py}: $|\alpha_{*}-2|<10^{-3}$
  & \textsf{NU}\\
17 & Cosmology: $\Delta\chi^{2}=+1.8$, $\Delta\mathrm{AIC}=+3.8$
  & vs.\ $\Lambda$CDM on Planck\,+\,DESI data, Friedmann~\eqref{eq:field-eq}
  & \textsf{NU}\\
18 & Solitonic $N$-body stability
  & Chaos suppression, Earnshaw violation; 10/10 tests pass
  & \textsf{NU}\\
19 & Metric derivation chain
  & Spatial $\to$ information $\to$ PPN; cross-checked with $\lP=\mathrm{const}$
  & \textsf{NU}\\
20 & Three-regimes existence
  & Numerical profile $\phi(r)$ with attraction, repulsion, confinement
  & \textsf{NU}\\
\bottomrule
\end{longtable}
\end{center}

\subsection{Open problems}

The following are explicitly open: they define the research frontier of
TGP at the core level.

\begin{center}\small
\begin{longtable}{@{}clp{8.5cm}c@{}}
\toprule
\# & \textbf{Open problem} & \textbf{Statement} & \textbf{Status}\\
\midrule
\endhead
OP-1 & Uniqueness of $U(\varphi)$
  & Is the cubic+quartic form the unique self-interference potential
  compatible with $\Zp$ symmetry and the vacuum condition?
  & \textsf{HY}\\
OP-2 & Parametric reduction $\beta=\gamma$
  & Formal proof (beyond the vacuum condition) that $\beta=\gamma$ is
  protected at all scales.
  & \textsf{HY}\\
OP-3 & Identification $a_{\Gamma}=1/\PhiZero$
  & Whether the two fitted parameters coincide, reducing the free input
  count to one.
  & \textsf{HY}\\
OP-4 & Derivation of the quartic coefficient $\gamma$
  & An \textit{a priori} determination of $\gamma$ (equivalently
  $\Lambda_{E}$) from $H_{\Gamma}$ alone.
  & \textsf{HY}\\
OP-5 & Full non-linear $N$-body regime
  & Existence/uniqueness of solutions for strongly overlapping sources
  beyond perturbation theory.
  & \textsf{HY}\\
OP-6 & Substrate metric continuum limit
  & Rigorous Migdal--Kadanoff convergence of $H_{\Gamma}$ to the continuum
  action~\eqref{eq:S-TGP}.
  & \textsf{HY}\\
OP-7 & Tensor projection $\sigma_{ab}$
  & Derivation of the explicit tensor-field equation from the same
  $H_{\Gamma}$; GW polarisation content.
  & \textsf{HY}\\
OP-8 & Gauge-sector emergence
  & Emergence of U(1)$\times$SU(2)$\times$SU(3) from the substrate
  (addressed partially in the extended manuscript).
  & \textsf{HY}\\
OP-9 & Strong-field black-hole solutions
  & Complete catalogue of non-singular BH solutions; quasi-normal modes.
  & \textsf{HY}\\
OP-10 & Quantum theory of $\Phi$
  & Second quantisation of $\Phi$ beyond the MS linear regime.
  & \textsf{HY}\\
\bottomrule
\end{longtable}
\end{center}

% =============================================================================
\section{Applications and falsifiability}\label{sec:applications}
% =============================================================================

The minimal core of TGP defined above supports several independent
applications. Each is developed in a dedicated companion paper that cites
this reference.

\begin{itemize}[nosep]
  \item \textbf{Superconductivity.} The substrate harmonic
    $a^{\ast}_{\mathrm{TGP}}=7.725\,a_{\mathrm{B}}\approx 4.088$\,\AA,
    together with the orbital-amplitude spectrum
    $\{A_{s},A_{sp},A_{d},A_{f}\}$, yields a unified prediction of
    superconducting critical temperatures across cuprates, phonon-mediated
    metals, pressure-driven systems and lanthanide hydrides, using three
    universal TGP constants plus tabulated atomic inputs. See the companion
    paper \cite{TGP-SC}.
  \item \textbf{Lepton masses.} A single substrate parameter
    $g_{0}^{e}=0.86941$ reproduces $m_{\mu}/m_{e}$ and $m_{\tau}/m_{e}$
    with better than $10^{-3}$ accuracy, and the Koide combination
    $K=2/3$ follows from an arithmetic identity of the Brannen
    parametrisation \cite{TGP-leptons}.
  \item \textbf{Black-hole shadows.} The exponential
    metric~\eqref{eq:exp-metric} in the strong-field regime predicts a
    photon-ring radius reduced by $\sim 37\%$ relative to Schwarzschild,
    falsifiable with next-generation EHT data \cite{TGP-BH}.
  \item \textbf{$N$-body dynamics.} Chaos suppression and a violation of
    Earnshaw's theorem in $N$-body configurations dominated by the
    intermediate-repulsion regime (regime~ii), tested numerically against
    explicit integration \cite{TGP-nbody}.
\end{itemize}

\paragraph{Minimal falsification targets.} The following predictions are
of a binary character and provide the most direct falsification tests of
the core theory:
\begin{enumerate}[(F1),nosep]
  \item $\alpha=2$ \emph{identically} (Theorem~\ref{thm:alpha2}):
    any observable inconsistent with $K\propto\varphi^{4}$ at the
    substrate level falsifies TGP.
  \item $\gamma_{\mathrm{PPN}}=\beta_{\mathrm{PPN}}=1$ \emph{to all
    orders} (Corollary~\ref{cor:ppn}): a measured deviation at
    $O(U^{2}/c_{0}^{4})$ beyond GR would falsify the exponential metric.
  \item $c_{\mathrm{GW}}=c_{0}$ \emph{identically}
    (Corollary~\ref{cor:cGW}): already supported by GW170817 at the
    $10^{-15}$ level.
\end{enumerate}

% =============================================================================
\section{Conclusion}\label{sec:conclusion}
% =============================================================================

We have presented a minimal, self-contained core of the Theory of Generated
Space: four axioms, one scalar field with a geometrically fixed kinetic
operator ($\alpha=2$), a single non-linear potential with two coefficients
related by a vacuum condition, and an algebraically determined effective
metric reproducing general relativity exactly at the post-Newtonian level.
Fifteen core results are formally proven, five more are numerically
verified, and ten explicit open problems are stated so that the theory is
both well-defined and falsifiable on a clear basis.

This core is intended to serve as a common reference for applications to
superconductivity, particle masses, black-hole phenomenology and cosmology,
developed in companion papers.

\paragraph{Data and reproducibility.} All numerical scripts underlying the
results of this paper are available in the accompanying source repository
\cite{TGP-repo}. The present paper deposits a fixed, versioned snapshot on
Zenodo (DOI assigned at publication).

\paragraph{Acknowledgements.} The author thanks the community of
open-source physics for the tools that made this work possible.

% =============================================================================
\begin{thebibliography}{99}
% =============================================================================

\bibitem{TGP-repo}
M.~Serafin, \emph{TGP source repository (v1)},
\url{https://github.com/mserafin/tgp} (2026).

\bibitem{TGP-SC}
M.~Serafin,
\emph{Unified superconductivity from the TGP substrate},
companion paper, Zenodo (2026).

\bibitem{TGP-leptons}
M.~Serafin,
\emph{Lepton masses and the Koide relation from a single TGP parameter},
companion paper (2026).

\bibitem{TGP-BH}
M.~Serafin,
\emph{Black-hole shadows in the TGP exponential metric},
companion paper (2026).

\bibitem{TGP-nbody}
M.~Serafin,
\emph{N-body dynamics in TGP: chaos suppression and Earnshaw violation},
companion paper (2026).

\bibitem{Will2014}
C.~M.~Will,
\emph{The confrontation between general relativity and experiment},
Living Rev.\ Relativity \textbf{17}, 4 (2014).

\bibitem{Planck2018}
Planck Collaboration,
\emph{Planck 2018 results.\ VI.\ Cosmological parameters},
A\&A \textbf{641}, A6 (2020).

\bibitem{GW170817-speed}
LIGO/Virgo Collaboration,
\emph{Gravitational waves and gamma-rays from a binary neutron star merger:
GW170817 and GRB 170817A}, ApJL \textbf{848}, L13 (2017).

\end{thebibliography}

\end{document}
