% tgp_perturbations_lss.tex -- M2 deliverable: Cosmological perturbations and LSS
% Integrates into sek08_formalizm.tex (after ssec:perturb) or sek07_predykcje.tex
% Cross-refs: thm:Geff-k, cor:grav-slip, prop:ghost-free-MS, eq:Geff-k

\subsection{Large-scale structure predictions}%
\label{ssec:lss-predictions}

The scale-dependent effective gravitational constant
(tw.~\ref{thm:Geff-k}) and the gravitational slip
(wn.~\ref{cor:grav-slip}) yield concrete, falsifiable
predictions for large-scale structure observables.
This subsection quantifies the predictions and compares
them with existing data from BOSS, DESI, and forthcoming
surveys (Euclid, LSST, SKA).

% ----- Stability ----- %
\subsubsection{Stability analysis}%
\label{sssec:lss-stability}

\begin{proposition}[Complete perturbative stability]%
\label{prop:lss-stability}
Cosmological perturbations of the TGP scalar field $\Phi$
around the homogeneous FRW background $\bar\Phi(t)$
satisfy \textbf{all} stability conditions:
\begin{enumerate}[(i),nosep]
  \item \textbf{No-ghost}: the kinetic coefficient in the
    Mukhanov--Sasaki sector is
    $Q_s = \psi_{\rm bg}^4 > 0$ for all $\psi_{\rm bg} > 0$
    (stw.~\ref{prop:ghost-free-MS});
  \item \textbf{Gradient stability}: the sound speed is
    $c_s^2 = c_0^2 > 0$ (identically the speed of light);
  \item \textbf{Tachyonic stability}: the spatial Yukawa mass satisfies
    $m_{\mathrm{sp}}^2 = \gamma > 0$
    (stw.~\ref{prop:vacuum-stability}), ensuring exponential
    suppression of $\delta\Phi$ on sub-Compton scales;
  \item \textbf{Cosmological mass}: the homogeneous mass
    $m_{\rm cosmo}^2 = \tfrac{4}{3}\,c_0^2\gamma > 0$
    corresponds to damped oscillations of $\bar\Phi$
    around the vacuum~$\Phi_0$, driven to rest by
    Hubble friction --- this is the relaxation to
    $\Lambda$CDM, not an instability.
\end{enumerate}
\end{proposition}

\noindent\textit{Numerical verification:}
Script \texttt{tgp\_perturbations\_formal.py} confirms all four
conditions over the range $\psi \in (0, 3]$ (18/18 tests PASS).

% ----- Growth factor ----- %
\subsubsection{Linear growth factor and $f\sigma_8$}%
\label{sssec:growth-lss}

The linear density contrast $\delta_m$ obeys:
\begin{equation}\label{eq:growth-ode}
  D'' + \bigl(2 + \tfrac{H'}{H}\bigr)\,D'
  = \tfrac{3}{2}\,\Omega_m(a)\,
    \frac{G_{\mathrm{eff}}(k,a)}{G_0}\;D,
\end{equation}
where primes denote $d/d\ln a$. The scale dependence of
$G_{\mathrm{eff}}$ (r.~\eqref{eq:Geff-k}) induces, in principle,
a scale-dependent growth rate $f(k,z) = d\ln D / d\ln a$.

\begin{table}[h]
\centering\small
\caption{Linear growth factor and growth rate $f\sigma_8$:
TGP vs.\ $\Lambda$CDM at $k = 0.1\;\mathrm{Mpc}^{-1}$.
At $\Phi_0 \approx 25$, deviations are below $10^{-50}$.}
\label{tab:growth-comparison}
\begin{tabular}{@{}ccccc@{}}
\toprule
$z$ & $D_{\Lambda\mathrm{CDM}}$ & $D_{\mathrm{TGP}}$ &
$|\Delta D/D|$ & $f\sigma_8$ \\
\midrule
0.0 & 1.0000 & 1.0000 & $< 10^{-50}$ & 0.428 \\
0.3 & 0.8528 & 0.8528 & $< 10^{-50}$ & 0.474 \\
0.5 & 0.7689 & 0.7689 & $< 10^{-50}$ & 0.476 \\
0.7 & 0.6964 & 0.6964 & $< 10^{-50}$ & 0.462 \\
1.0 & 0.6067 & 0.6067 & $< 10^{-50}$ & 0.431 \\
1.5 & 0.4956 & 0.4956 & $< 10^{-50}$ & 0.374 \\
2.0 & 0.4172 & 0.4172 & $< 10^{-50}$ & 0.324 \\
\bottomrule
\end{tabular}
\end{table}

The growth-rate observable $f\sigma_8(z) = f(z)\,\sigma_8\,D(z)$
agrees with $\Lambda$CDM to machine precision at $\Phi_0 \approx 25$,
giving $\chi^2 = 9.86$ for 14 combined data points
(BOSS DR12 + 6dFGS + VIPERS + eBOSS + DESI DR1 RSD).

\begin{remark}[Observational bound on $\alpha_{\mathrm{eff}}$]%
\label{rem:alpha-bound}
A $\chi^2$ scan over $\alpha_{\mathrm{eff}}$ using the combined
$f\sigma_8$ compilation yields the $2\sigma$ upper bound:
\begin{equation}\label{eq:alpha-bound}
  \alpha_{\mathrm{eff}} < 0.17 \quad (2\sigma).
\end{equation}
The TGP prediction $\alpha_{\mathrm{eff}} = q\,\Phi_0/(4\pi)
\approx 3.7 \times 10^{-26}$ lies $25$~orders of magnitude below
this threshold. Scale-dependent growth would become detectable
only for $\alpha_{\mathrm{eff}} \gtrsim 0.01$, corresponding to
$\Phi_0 \gtrsim 10^{24}$.
\end{remark}

% ----- Modified gravity parameterization ----- %
\subsubsection{Modified gravity parameterization $(\mu, \Sigma, \eta)$}%
\label{sssec:mu-sigma-eta}

In the standard parameterized post-Friedmann framework:
\begin{align}
  \mu(k,a) &\equiv \frac{G_{\mathrm{eff}}}{G_0}
  = 1 + \frac{2\,\alpha_{\mathrm{eff}}^2}
        {1 + (a\,m_{\mathrm{sp}}/k)^2},
  \label{eq:mu-lss} \\[4pt]
  \Sigma(k,a) &= 1 + \frac{\alpha_{\mathrm{eff}}^2}
        {1 + (a\,m_{\mathrm{sp}}/k)^2},
  \label{eq:sigma-lss} \\[4pt]
  \eta(k,a) &\equiv \frac{\Sigma}{\mu}
  = \frac{1 + \alpha_{\mathrm{eff}}^2/(1+x^2)}
         {1 + 2\alpha_{\mathrm{eff}}^2/(1+x^2)},
  \qquad x = a\,m_{\mathrm{sp}}/k.
  \label{eq:eta-lss}
\end{align}
For $k/a \gg m_{\mathrm{sp}}$ (small scales):
$\eta \approx 1 - \alpha_{\mathrm{eff}}^2$.
For $k/a \ll m_{\mathrm{sp}}$ (large scales):
$\eta \to 1$ (GR recovered).

At $\Phi_0 \approx 25$:
$\mu - 1 \sim \Sigma - 1 \sim \eta - 1 \sim 10^{-51}$,
far below the projected sensitivity of
Euclid ($\sigma_\mu \sim 0.02$, Amendola et al.\ 2018)
and LSST ($\sigma_\Sigma \sim 0.03$).

\begin{table}[h]
\centering\small
\caption{Modified gravity parameters at
$k = 0.1\;\mathrm{Mpc}^{-1}$ for TGP ($\Phi_0 \approx 25$).
All deviations from GR are undetectable.}
\label{tab:mu-sigma-eta}
\begin{tabular}{@{}cccc@{}}
\toprule
$z$ & $\mu - 1$ & $\Sigma - 1$ & $\eta - 1$ \\
\midrule
0.0 & $< 10^{-51}$ & $< 10^{-51}$ & $< 10^{-51}$ \\
0.5 & $< 10^{-51}$ & $< 10^{-51}$ & $< 10^{-51}$ \\
1.0 & $< 10^{-51}$ & $< 10^{-51}$ & $< 10^{-51}$ \\
2.0 & $< 10^{-51}$ & $< 10^{-51}$ & $< 10^{-51}$ \\
\bottomrule
\end{tabular}
\end{table}

% ----- Inflationary sector ----- %
\subsubsection{Inflationary predictions}%
\label{sssec:inflation-lss}

The Mukhanov--Sasaki equation in TGP reads (prop.~\ref{prop:ghost-free-MS}):
\begin{equation}\label{eq:MS-TGP}
  v_k'' + \bigl[c_s^2\,k^2 - z''/z\bigr]\,v_k = 0,
  \qquad z = a\,\psi_{\rm bg}^2,\quad c_s^2 = c_0^2.
\end{equation}
In the slow-roll regime (Dodatek~G), the spectral index and
tensor-to-scalar ratio are:
\begin{equation}\label{eq:ns-r-TGP}
  n_s = 1 - \frac{2}{N_e}, \qquad
  r = \frac{16}{N_e}, \qquad
  f_{\rm NL}^{\rm local} \approx -\frac{1}{N_e},
\end{equation}
where $N_e$ is the number of $e$-folds.
For $N_e \in [55, 65]$:
$n_s \in [0.964, 0.969]$ (Planck $1\sigma$),
$r \in [0.25, 0.29]$,\footnote{%
  The standard relation $r = 16\varepsilon$ assumes
  an unmodified tensor sector. In TGP, the tensor sector
  ($\sigma_{ab}$) has status Szkic; the actual $r_{\rm TGP}$
  may differ. See sek.~\ref{sssec:sigma-status-map}.}
$|f_{\rm NL}| \lesssim 0.02$ (undetectable by Planck).

% ----- Falsifiability ----- %
\subsubsection{Falsifiability criteria}%
\label{sssec:falsifiability-lss}

\begin{table}[h]
\centering\small
\caption{Falsifiability of TGP in the perturbation sector.}
\label{tab:falsifiability}
\begin{tabular}{@{}lp{5.5cm}l@{}}
\toprule
\textbf{Observable} & \textbf{Condition to falsify TGP} &
\textbf{Survey} \\
\midrule
$\mu(k,z)$ & $|\mu - 1| > 0.02$ at fixed $\Phi_0 \approx 25$
& Euclid \\
$\Sigma(k,z)$ & $|\Sigma - 1| > 0.03$
& LSST \\
$f\sigma_8(z)$ & Scale-dependent anomaly at
$\Delta(f\sigma_8)/f\sigma_8 > 1\%$
& DESI, Euclid \\
$n_s$ & $n_s \notin [0.95, 0.98]$ (no viable $N_e$)
& CMB-S4 \\
$c_{\rm GW}$ & $|c_{\rm GW}/c_0 - 1| > 10^{-15}$
& LIGO/ET \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}[Perturbation sector conclusion]%
\label{rem:perturbation-conclusion}
At $\Phi_0 \approx 25$ (the value fixed by
$\Lambda_{\mathrm{eff}} \approx \Lambda_{\mathrm{obs}}$),
TGP is \textbf{operationally equivalent to $\Lambda$CDM}
in the cosmological perturbation sector:
$G_{\mathrm{eff}} = G_0$ to $O(10^{-51})$,
all stability conditions are satisfied,
and the inflationary predictions lie within the
Planck $1\sigma$ band for $N_e \sim 57$--$65$.

The distinct predictions of TGP lie
\textbf{outside} the linear perturbation regime:
\begin{enumerate}[(a),nosep]
  \item modified dark energy equation of state $w_{\rm DE}(z)$
    (\S\ref{ssec:de-oszacowanie}, likelihood H6:
    $\Delta\chi^2 = +1.8$ vs.\ $\Lambda$CDM);
  \item GW dispersion at the Yukawa mass scale
    $m_{\mathrm{sp}} \sim 10^{-26}\;\mathrm{m}^{-1}$
    (Table~\ref{tab:gw-summary});
  \item $N$-body dynamics: Earnshaw violation, chaos
    suppression, three-body force (Appendix~\ref{app:nbody}).
\end{enumerate}
These are the arenas where TGP can be distinguished from GR.
\end{remark}
