% ============================================================
% sec08_predictions.tex — Predictions, falsifiability, N-body
% ============================================================

\section{Predictions and falsifiability}\label{sec:predictions}

% ==================================================================
\subsection{Particle sector: the strongest argument for TGP}%
\label{ssec:particle-sector}
% ==================================================================

TGP is not merely a theory of gravity: as a theory of the scalar
substrate from which the metric, fundamental constants, \emph{and}
particle masses emerge, TGP possesses a particle sector yielding
quantitative predictions unmatched by any other modified-gravity
theory.

\subsubsection{Lepton masses from one parameter}

The substrate soliton ODE
\begin{equation}\label{eq:soliton-ode}
  g^{2}\,g'' + g\,(g')^{2} + \frac{2}{r}\,g^{2}\,g' = g^{2}(1-g)
\end{equation}
has a one-parameter family of solutions labeled by the central
value $g_{0}$.  A single calibration parameter $g_{0}^{e}=0.869$
(fixed by the muon-to-electron mass ratio $r_{21}=m_{\mu}/m_{e}$
via the $\varphi$-fixed-point mechanism) yields:

\begin{center}\small
\begin{tabular}{@{}llrl@{}}
\toprule
\textbf{Observable} & \textbf{TGP prediction} &
\textbf{PDG value}~\cite{PDG2024} & \textbf{Accuracy} \\
\midrule
$r_{21} = m_{\mu}/m_{e}$
  & $\bigl[A_{\mathrm{tail}}(\varphi\,g_{0}^{e})/A_{\mathrm{tail}}(g_{0}^{e})\bigr]^{4} = 206.77$
  & 206.768
  & 0.001\% \\[4pt]
$m_{\tau}$ (via Koide $K=2/3$)
  & 1776.9\,MeV
  & 1776.86\,MeV
  & 0.002\% \\
\bottomrule
\end{tabular}
\end{center}

\noindent
Seven additional lepton mass ratios are reproduced at
$<0.1\%$ precision from the same $g_{0}^{e}$.

\subsubsection{Strong coupling constant: zero free parameters}

The same $g_{0}^{e}$ (determined purely from the lepton sector)
predicts the QCD coupling at the $Z$-boson mass:
\begin{equation}\label{eq:alphas}
  \alpha_{s}(M_{Z})
  = \frac{N_{c}^{3}\,g_{0}^{e}}{8\,N_{f}^{2}}
  = \frac{27 \times 0.869}{8 \times 25}
  = 0.1174
  \qquad (\text{PDG: } 0.1179 \pm 0.0009;\; 0.6\sigma).
\end{equation}
This is a prediction with \textbf{zero} free parameters from the
QCD or cosmological sectors; $g_{0}^{e}$ comes entirely from lepton
masses.

\subsubsection{Discrete running: parameter-free test}

The ratio of $\alpha_{s}$ at two flavor thresholds
($N_{f}=3$ at $m_{\tau}$; $N_{f}=5$ at $M_{Z}$) is a
\textbf{parameter-free} prediction:
\begin{equation}\label{eq:running-ratio}
  \frac{\alpha_{s}(m_{\tau})}{\alpha_{s}(M_{Z})}
  = \left(\frac{5}{3}\right)^{2}
  = \frac{25}{9}
  = 2.778
  \qquad (\text{PDG: } 2.799 \pm 0.121;\; 0.18\sigma).
\end{equation}

\begin{remark}[Implication]\label{rem:lepton-qcd}
The chain $r_{21}\to g_{0}^{e}\to\alpha_{s}(M_{Z})$ constitutes
a closed map: from one electroweak observable (lepton mass ratio),
a QCD observable (strong coupling) follows.  This is a feature
unique to TGP, impossible in purely gravitational theories.
\end{remark}

% ==================================================================
\subsection{Kill-shot tests (falsifiability)}\label{ssec:killshots}
% ==================================================================

TGP identifies four sharp falsification criteria, each targeting
a structural prediction:

\begin{center}\small
\begin{tabular}{@{}clp{5.5cm}l@{}}
\toprule
\textbf{ID} & \textbf{Observable} & \textbf{Falsification condition} &
\textbf{Status} \\
\midrule
K-A & $c_{\mathrm{GW}}$
  & $c_{\mathrm{GW}} \neq c_{0}$ at any level
  & \textbf{PASSED} (GW170817: $|c_{\mathrm{GW}}/c-1|<3\times 10^{-15}$) \\[4pt]
K-B & $\gamma_{\mathrm{PPN}}$
  & $\gamma_{\mathrm{PPN}} \neq 1$ at any level
  & \textbf{PASSED} (Cassini: $|\gamma-1|<2.3\times 10^{-5}$) \\[4pt]
K-C & $G_{\mathrm{eff}}(k)$
  & Scale-dependent $G_{\mathrm{eff}}$ at detectable level
  & \textbf{PASSED} ($\alpha_{\mathrm{eff}}\sim 10^{-26}$; undetectable) \\[4pt]
K-D & Lepton mass ratios
  & $r_{21}$ or $m_{\tau}$ deviate from substrate ODE predictions
  & \textbf{PASSED} (0.001\% and 0.002\%) \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Future kill-shots.}
\begin{itemize}[nosep]
  \item \textbf{K-E}: Phantom crossing $w_{\mathrm{DE}}(z)<-1$
        detected at any redshift $\Rightarrow$ TGP falsified
        (the potential enforces $w_{\mathrm{DE}}\geq -1$).
  \item \textbf{K-F}: Detection of $|\mu-1|>0.02$ or
        $|\Sigma-1|>0.03$ at fixed $\PhiZero\approx 25$
        (Euclid, LSST).
  \item \textbf{K-G}: $n_{s}\notin[0.95,0.98]$ with no viable
        $N_{e}$ (CMB-S4).
  \item \textbf{K-H}: Detection of sharp photon-ring substructure
        at $R_{\mathrm{ph}} = 3GM/c^{2}$ consistent with GR
        (ngEHT) $\Rightarrow$ TGP falsified.  The soliton
        $\psi = (1+3r_{s}/r)^{1/3}$ has no photon sphere; the
        $1/3$ exponent structurally forbids it for \emph{any}
        lattice geometry (harmonicity protection, discrete
        Green's function regularization).
\end{itemize}

% ==================================================================
\subsection{N-body dynamics: summary of 8 Results}%
\label{ssec:nbody-summary}
% ==================================================================

\noindent\textit{(Full analysis: \texttt{dodatekY\_nbody.tex} and
\texttt{nbody/tgp\_nbody\_results\_clean.tex}; 59/59 regression
tests PASS, 102 active scripts.)}

\medskip
The $N$-body sector arises from the TGP scalar-field Lagrangian
with quartic self-interaction $V(\varphi)
=(\beta/3)\varphi^{3}-(\gamma/4)\varphi^{4}$.  The pairwise
potential $V_{2}$ exhibits the three regimes of TGP (attraction,
repulsion, confinement); the irreducible three-body force $V_{3}$
arises from the $\Phi^{4}$ vertex.  Eight principal results are:

\begin{enumerate}[nosep]
  \item \textbf{Earnshaw violation.}
        Equilateral 3-body configurations admit \emph{stable static
        equilibria} with all normal modes $\omega^{2}>0$.
        In Newtonian gravity, Earnshaw's theorem forbids such
        equilibria; TGP evades it via the short-range repulsive
        barrier.

  \item \textbf{Chaos suppression (35--45\%).}
        The maximal Lyapunov exponent of the 3-body problem is
        reduced by 35--45\% for $\beta\gtrsim 0.025$, enhanced
        by the three-body force $V_{3}$.  The repulsive barrier
        prevents close encounters that drive chaotic divergence
        in Newtonian dynamics.

  \item \textbf{$C_{\mathrm{eff}}$ from EFT.}
        The Yukawa source coupling $C_{\mathrm{eff}}$ is derived
        from an EFT projection of the classical topological defect
        (soliton) profile, connecting the microscopic substrate
        to the macroscopic $N$-body force law.

  \item \textbf{Classical $\neq$ Yukawa.}
        The classical soliton--soliton interaction is
        \emph{oscillatory} with power-law decay, qualitatively
        different from the Yukawa law $\propto e^{-m_{\mathrm{sp}}r}/r$
        that requires the EFT mass gap.

  \item \textbf{$I_{Y}$ multipole representation.}
        The irreducible three-body Feynman integral $I_{Y}$ admits
        a semi-analytic multipole expansion converging to
        $<0.01\%$ at $L_{\max}=15$.

  \item \textbf{$V_{3}/V_{2}$ regime map.}
        The three-body force is perturbative ($\sim 1\%$) in the
        Lyapunov regime; exponentially suppressed for
        $m_{\mathrm{sp}}\,d > 3$.

  \item \textbf{Analytical equilibria and Hill sphere.}
        Closed-form equilibrium distances $d_{\mathrm{rep}}$,
        $d_{\mathrm{well}}$; radial normal-mode frequency;
        TGP Hill sphere generically \emph{expanded} relative to
        Newton.

  \item \textbf{Unequal-mass phase diagram.}
        Phase boundary
        $\beta_{\mathrm{crit}}=\frac{3}{2}\sqrt{\gamma(C_{1}+C_{2})}$;
        escape velocity
        $v_{\mathrm{esc}}=\sqrt{2|V_{2}(d_{\mathrm{well}})|/\mu}$;
        breathing mode frequency
        $\omega_{\mathrm{br}}^{2}=\frac{3}{2}\,\omega_{\mathrm{rad}}^{2}$
        (exact), with $V_{3}$ correction of $-1.3\%$.
\end{enumerate}

\begin{remark}[Relation to main formalism]\label{rem:nbody-connection}
The pairwise potential $V_{2}$ (gradient + repulsion + confinement)
maps directly onto the three force regimes of TGP
(\S\ref{sec:formalism}).  The three-body force $V_{3}$ arises from
the $\Phi^{4}$ self-interaction in the unified action
(eq.~\ref{eq:S-TGP}).  The EFT coupling $C_{\mathrm{eff}}$
provides the microscopic origin of the source strength $C_{i}$
used throughout the $N$-body analysis, connecting to the soliton
mass hierarchy ($r_{21}=206.77$) of the particle sector.
\end{remark}

% ==================================================================
\subsection{Where TGP departs from GR}\label{ssec:departures}
% ==================================================================

The PPN identity and the $\Lambda$CDM equivalence in the LSS sector
do \emph{not} mean TGP\,$=$\,GR.
Deviations appear in five arenas:

\begin{enumerate}[nosep]
  \item \textbf{Cosmological dark energy:}
        Modified Friedmann equation with $G(\Phi)$ and
        $\Lambda_{\mathrm{eff}}=\gamma/12$; time-varying
        $w_{\mathrm{DE}}(z)$ controlled by $\PhiZero$
        ($\Delta\chi^{2}=+1.8$ vs.\ $\Lambda$CDM).

  \item \textbf{Strong-field regime:}
        Frozen interior (no singularity) replacing the
        Schwarzschild/Kerr singularity.  The substitution
        $u = \psi^{3}$ linearises the kinetic operator to
        $\nabla^{2}u = 0$, yielding the exact BH soliton
        $\psi(r) = (1 + 3r_{s}/r)^{1/3}$.  This solution has
        $g_{tt}\to 0$ at the centre but \textbf{no photon sphere}:
        $h(r) = r\,\psi(r)$ is strictly increasing
        (companion paper~\cite{TGP-repo}).

  \item \textbf{Scalar breathing mode:}
        Massive ($m_{\mathrm{sp}}\sim H_{0}$), screened in the
        LIGO band, detectable in the PTA band
        ($h_{\mathrm{br}}\sim 10^{-17}$) with a monopolar angular
        pattern.

  \item \textbf{Three-body forces:}
        $V_{3}$ from $\Phi^{4}$ self-interaction, with no analogue
        in GR; Earnshaw violation, chaos suppression, expanded
        Hill sphere.

  \item \textbf{Particle physics:}
        Lepton mass hierarchy from soliton ODE
        ($r_{21}=206.77$, $m_{\tau}=1776.9$\,MeV);
        $\alpha_{s}(M_{Z})=0.1174$ with zero free parameters.
\end{enumerate}

\noindent
These are the sectors where TGP can be \emph{distinguished} from
GR and $\Lambda$CDM by present or near-future experiments.
