% tgp_metrology_observables.tex -- Dimensionless observables and metrological protocol
% Self-contained section for \input{} into main manuscript.
% References: BIPM 2019 (SI redefinition), CODATA 2018/2022.

\subsection{Dimensionless observables and metrological consistency}%
\label{ssec:metrology}

\paragraph{The problem.}
TGP posits $c(\Phi)$, $\hbar(\Phi)$, $G(\Phi)$ as field-dependent.
Since the 2019 SI fixes $c$ and $h$ as \emph{exact constants}
(BIPM, Resolution~1 of the 26th CGPM), any claim that these
quantities ``vary'' must be stated carefully.

\paragraph{Resolution.}
The physical content of TGP is fully expressed through
\textbf{dimensionless combinations} that are convention-independent:

\begin{center}\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Observable} & \textbf{Expression} & \textbf{TGP dependence on $\Phi$} & \textbf{Test} \\
\midrule
Fine structure constant $\alpha$
  & $e^2/(4\pi\varepsilon_0\hbar c)$
  & $\alpha(\Phi) = \alpha_0$ \quad (\textbf{constant})
  & Atomic clocks; quasar absorption \\[4pt]
Proton-to-electron mass ratio $\mu$
  & $m_p/m_e$
  & $\mu(\Phi) = \mu_0$ \quad (\textbf{constant})
  & Molecular spectroscopy \\[4pt]
Gravitational fine structure $\alpha_G$
  & $G m_p^2/(\hbar c)$
  & $\alpha_G(\Phi) = \alpha_{G,0}\,/\psi^2$ \quad (\textbf{varies})
  & Binary pulsar timing \\[4pt]
Gravitational redshift $z_{\rm grav}$
  & $\nu_A/\nu_B - 1$
  & $z_{\rm grav} = \sqrt{\psi_B/\psi_A} - 1$
  & Gravity Probe A; ACES; GPS \\[4pt]
Hubble rate $H(z)$
  & $\dot{a}/a$
  & $H(z;\Phi_0,\gamma)$ (modified Friedmann)
  & BAO (DESI); CMB (Planck) \\[4pt]
GW/EM speed ratio
  & $c_{\rm GW}/c_{\rm EM}$
  & $= 1$ \quad (exact, algebraic identity)
  & Multi-messenger (GW170817) \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Key point.}
In TGP, $\alpha$ and $\mu$ are \textbf{$\Phi$-independent}
because both $e^2/(\hbar c)$ and $m_p/m_e$ involve ratios
where the $\Phi$-dependence cancels.  This is a non-trivial
consistency condition: TGP predicts that atomic physics is
locally invariant even though individual ``constants'' vary.
Only \emph{gravitational} dimensionless quantities ($\alpha_G$,
$z_{\rm grav}$) carry $\Phi$-dependence.

\paragraph{CODATA protocol.}
All numerical comparisons in this work use:
\begin{itemize}[nosep]
  \item CODATA 2018 recommended values for fundamental constants
        (NIST SP~961, 2019 edition);
  \item PDG 2024 for particle masses and coupling constants;
  \item Planck 2018 (VI, Table~2) for cosmological parameters
        ($H_0 = 67.36 \pm 0.54$\,km/s/Mpc, $\Omega_m = 0.3153$, etc.);
  \item DESI DR1 BAO distances for $H(z)$ and $D_M(z)$ constraints
        (DESI Collaboration 2024, arXiv:2404.03002).
\end{itemize}
When TGP and GR predictions differ, the comparison is always
stated as a deviation in a \textbf{dimensionless observable}
(not in a dimensional ``constant'').

\paragraph{Experimental sensitivity.}
Current bounds on variation of $\alpha$ and $\mu$:
$|\dot\alpha/\alpha| < 10^{-17}\,{\rm yr}^{-1}$ (optical clocks, PTB/NIST),
$|\dot\mu/\mu| < 10^{-16}\,{\rm yr}^{-1}$ (molecular spectroscopy).
TGP predicts $\dot\alpha = \dot\mu = 0$ at all $\Phi$ ---
these bounds are \textbf{automatically satisfied}.
The non-trivial TGP signal is in $\alpha_G(\Phi)$: the effective
gravitational coupling varies as $G_{\rm eff} \propto 1/\psi^2$.
For $\psi = 1 + 2U/c^2$, the deviation is
$\delta G/G_0 \approx -4U/c^2 \sim -6 \times 10^{-10}$
on the surface of the Sun --- consistent with helioseismology
bounds ($|\delta G/G| < 10^{-4}$, Guenther et al.\ 1998).
