% tgp_lensing_geff.tex -- M3 deliverable: Gravitational lensing in TGP
% Integrates after sssec:lensing in sek08_formalizm.tex
% Extends qualitative predictions from sek07 with full analytical derivation

\subsubsection{Analytical lensing in $g_{\mathrm{eff}}$:
  PPN connection and Yukawa corrections}%
\label{sssec:lensing-analytical}

The preceding subsection (\S\ref{sssec:lensing}) established the
deflection angle and lensing convergence \emph{qualitatively}.
Here we close the analytical derivation, connecting:
(i)~the point-mass lensing to PPN parameters ($\gamma_{\rm PPN} = 1$),
(ii)~the Yukawa correction from the scalar propagator $\delta\Phi$,
(iii)~the Shapiro time delay (Cassini bound), and
(iv)~the post-post-Newtonian (ppN) regime where TGP's exponential
metric diverges from the Schwarzschild solution.

% --- Born integral via angular substitution ---
\paragraph{Exact Born integral for exponential metric.}
For the TGP metric
$ds^2 = -(c_0^2/\psi)\,dt^2 + \psi\,\delta_{ij}\,dx^i dx^j$
with $\psi = e^{2U/c_0^2}$, $U = GM/r$ (unsigned),
the effective refractive index is
$n(r) = \psi(r) = e^{2U/c_0^2}$.
The deflection angle in the Born approximation reads
\begin{equation}\label{eq:born-integral}
  \hat\alpha
  = -\int_{-\infty}^{+\infty}
    \nabla_\perp \!\ln\psi\;dl
  = \int_{-\infty}^{+\infty}
    \frac{2\,G\,M\,b}{c_0^2\,(b^2 + l^2)^{3/2}}\;dl,
\end{equation}
where $b$ is the impact parameter and $l$ the coordinate along
the unperturbed ray ($r^2 = b^2 + l^2$).
Substituting $l = b\tan\theta$ collapses the integral to
\begin{equation}\label{eq:deflection-exact}
  \hat\alpha
  = \frac{2\,G\,M}{c_0^2\,b}
    \int_{-\pi/2}^{\pi/2}\!\cos\theta\;d\theta
  = \frac{4\,G\,M}{c_0^2\,b},
\end{equation}
\textbf{identically} the GR result. The key feature is that
$\ln\psi = 2U/c_0^2$ is \emph{exact} (not a weak-field expansion),
so the Born integral closes analytically at all orders in~$U$.
Numerical integration confirms eq.~\eqref{eq:deflection-exact}
to $< 10^{-6}$ relative precision
(skrypt \texttt{tgp\_lensing\_formal.py}, 8/8 PASS).

% --- PPN consistency ---
\paragraph{PPN consistency.}
The general PPN deflection angle is
$\hat\alpha = \tfrac{1}{2}(1+\gamma_{\rm PPN})\times 4GM/(c_0^2 b)$.
With $\gamma_{\rm PPN} = 1$ (eq.~\eqref{eq:gamma-beta}),
this reproduces eq.~\eqref{eq:deflection-exact}. Together with
$\beta_{\rm PPN} = 1$, $\alpha_{1,2} = 0$
(eq.~\eqref{eq:alpha12}), TGP passes all solar-system lensing
and timing tests identically to GR.

% --- Shapiro delay ---
\paragraph{Shapiro time delay.}
A signal traversing the gravitational field of mass~$M$
with closest approach~$b$ acquires a delay:
\begin{equation}\label{eq:shapiro-TGP}
  \Delta t = \frac{1}{c_0}\int\!\bigl[n(r) - 1\bigr]\,dl
           = \frac{1}{c_0}\int\!\bigl[e^{2U/(c_0^2 r)} - 1\bigr]\,dl
           \approx \frac{4\,G\,M}{c_0^3}\,\ln\!\frac{4r_1 r_2}{b^2},
\end{equation}
where $r_{1,2}$ are the emitter and receiver distances.
The linearization error is $O(U^2/c_0^4) \sim 10^{-12}$
(verified numerically at the Cassini configuration:
$|\Delta t_{\rm exact}/\Delta t_{\rm weak} - 1| = 3 \times 10^{-7}$).
The Cassini constraint $|\gamma - 1| = (2.1 \pm 2.3)\times 10^{-5}$
(Bertotti et al.\ 2003) is satisfied automatically:
TGP gives $\gamma - 1 = 0$ exactly.

% --- Yukawa correction ---
\paragraph{Yukawa correction from scalar propagator.}
The scalar perturbation $\delta\Phi$ around a point mass
satisfies $(\nabla^2 - m_{\rm sp}^2)\,\delta\Phi = -q\,\Phi_0\,\rho$,
yielding a Yukawa profile:
\begin{equation}\label{eq:yukawa-lens}
  \delta\psi_{\rm Yuk}(r)
  = 2\,\alpha_{\rm eff}^2\,\frac{e^{-m_{\rm sp}\,r}}{m_{\rm sp}\,r},
  \qquad
  \alpha_{\rm eff} = \frac{q\,\Phi_0}{4\pi}
  \approx 3.7 \times 10^{-26}.
\end{equation}
The Compton wavelength $\lambda_{\rm sp} = 2\pi/m_{\rm sp}
\approx 5.6\;\mathrm{Gpc}$ exceeds all astrophysical
scales, so the Yukawa factor is \emph{unsuppressed}
($m_{\rm sp} r \ll 1$ at solar, galactic, and cluster scales).
However, the amplitude is $\alpha_{\rm eff}^2 \sim 10^{-51}$:
\begin{equation}\label{eq:delta-alpha-yukawa}
  \frac{\Delta\hat\alpha_{\rm Yuk}}{\hat\alpha_{\rm GR}}
  \sim \alpha_{\rm eff}^2 \sim 10^{-51},
\end{equation}
which is $\sim\!49$ orders of magnitude below the Cassini bound.

% --- ppN correction ---
\paragraph{Post-post-Newtonian (ppN) correction.}
The exponential metric deviates from Schwarzschild at $O(U^2)$:
\begin{align}
  g_{rr}^{\rm TGP}  &= 1 + 2U/c_0^2 + 2U^2/c_0^4 + \ldots,
  \nonumber\\
  g_{rr}^{\rm Schw} &= 1 + 2U/c_0^2 + 4U^2/c_0^4 + \ldots.
  \label{eq:ppN-grr}
\end{align}
The Born integral (eq.~\ref{eq:born-integral}) is insensitive to
this difference because $\ln\psi$ is an exact logarithm;
the ppN correction enters only through \emph{ray-tracing}
(non-Born) effects. At the solar limb:
\begin{equation}\label{eq:ppN-estimate}
  \frac{\Delta\hat\alpha_{\rm ppN}}{\hat\alpha_{\rm GR}}
  \sim \left(\frac{G\,M_\odot}{R_\odot\,c_0^2}\right)^{\!2}
  \approx 4.5 \times 10^{-12},
\end{equation}
corresponding to $\sim\!8\;\mathrm{pico\text{-}arcsec}$
--- far below astrometric precision for decades to come.

% --- Strong-field predictions ---
\paragraph{Strong-field predictions (indicative).}
Extrapolating the exponential metric to the strong-field regime
yields qualitatively different predictions from Schwarzschild:

\begin{center}\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Observable} & \textbf{GR (Schwarzschild)} &
\textbf{TGP (exponential)} & $\Delta/\mathrm{GR}$ \\
\midrule
Event horizon   & $r_H = r_s = 2GM/c_0^2$  & brak (exp. $> 0$ zawsze) & --- \\[2pt]
Photon sphere   & $r_{\rm ph} = \tfrac{3}{2}\,r_s$ & $r_{\rm ph} = r_s$ & $-33\%$ \\[2pt]
Shadow $b_{\rm crit}$ & $3\sqrt{3}\;GM/c_0^2$ & $r_s\sqrt{e}$ & $-37\%$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent
\textbf{Caveat.}
These strong-field numbers assume the exponential metric
$\psi = e^{2U/c_0^2}$ extends to $U/c_0^2 \sim 0.5$.
In TGP, the actual strong-field solution may differ
(see \S\ref{sec:czarne} for the TGP black-hole model,
where the interior approaches the $\Nzero$ limit).
A~complete strong-field lensing calculation requires the
full radial profile $\psi(r)$ from the TGP field equation
--- this remains an open problem.

% --- Summary table ---
\begin{table}[h]
\centering\small
\caption{Gravitational lensing: TGP vs.\ GR summary.}
\label{tab:lensing-summary}
\begin{tabular}{@{}lp{4.5cm}p{3.5cm}@{}}
\toprule
\textbf{Test} & \textbf{TGP prediction} & \textbf{Status} \\
\midrule
Solar deflection   & $\hat\alpha = 4GM_\odot/(c_0^2 R_\odot)$
  = $1.7513''$ & $\equiv$ GR \\[3pt]
Shapiro delay      & $\gamma = 1$ exactly
  & Cassini: $\checkmark$ \\[3pt]
Yukawa correction  & $\Delta\hat\alpha/\hat\alpha \sim 10^{-51}$
  & Undetectable \\[3pt]
ppN correction     & $\Delta\hat\alpha \sim 8$ pico-arcsec
  & Undetectable \\[3pt]
Cosmic shear       & $\Sigma = 1 + O(10^{-51})$
  & $\equiv$ $\Lambda$CDM \\[3pt]
BH shadow          & $\Delta b_{\rm crit} \sim -37\%$
  & Open (needs full soln.) \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}[Lensing sector conclusion]\label{rem:lensing-conclusion}
In the weak-field regime, TGP's lensing predictions are
\textbf{identical to GR} --- this is a structural consequence
of the exponential metric and $\gamma_{\rm PPN} = 1$.
The Yukawa correction ($\alpha_{\rm eff}^2 \sim 10^{-51}$)
and ppN correction ($\sim 10^{-12}$) are both unobservable.

The \textbf{only} distinctive TGP lensing prediction lies
in the \emph{strong-field} regime: the exponential metric
never develops an event horizon, the photon sphere is smaller
by 33\%, and the shadow is smaller by 37\%.
These are in principle testable by the Event Horizon Telescope,
but a rigorous comparison requires the full TGP compact-object
solution (status: open problem, \S\ref{sec:czarne}).
\end{remark}
