\section{Efimov-like three-body bound states in TGP}
\label{sec:efimov}

\subsection{The question}

Can three-body TGP forces create a \emph{bound triangular state} that
does not exist in the two-body theory alone?  Specifically: is there a
regime where
\begin{equation}
  E_{\rm min}^{(2B+{\rm ZP})} > 0 \quad\text{(unbound pair)}
  \qquad\text{but}\qquad
  E_{\rm min}^{(2B+3B+{\rm ZP})} < 0 \quad\text{(bound triple)}?
\end{equation}
The effective energy including zero-point pressure is
\begin{equation}
  E_{\rm eff}(d;C,m_{\rm sp}) = \underbrace{\frac{3}{8d^2}}_{\rm ZP}
  + \underbrace{V_2(d;C,m_{\rm sp})}_{2\text{-body Yukawa}}
  + \underbrace{V_3(d;C,m_{\rm sp})}_{3\text{-body Feynman}},
\end{equation}
where $d$ is the equilateral triangle side length and
$V_3 = -C^3 I_Y(d,d,d;\,m_{\rm sp})$.

\subsection{Critical couplings}

Minimising $E_{\rm eff}$ over $d$ defines
$E_{\rm min}(C,m_{\rm sp})$.  The \textbf{critical coupling}
$C_{\rm crit}(m_{\rm sp})$ is where $E_{\rm min}=0$:
above this value the triangle is quantum-mechanically bound.

\begin{table}[h]
\centering
\caption{Critical couplings for two-body ($C_{\rm crit}^{(2B)}$) and
two-body+three-body ($C_{\rm crit}^{(3B)}$) binding vs scalar mass
$m_{\rm sp}$.  The \emph{window} column gives the interval of $C$ for
which the triple is bound but no pair is bound.}
\label{tab:Ccrit_window}
\begin{tabular}{rcccc}
\hline
$m_{\rm sp}$ [$l_{\rm Pl}^{-1}$] & $\lambda=1/m_{\rm sp}$ [$l_{\rm Pl}$]
  & $C_{\rm crit}^{(2B)}$ & $C_{\rm crit}^{(3B)}$ & Window width \\
\hline
$0.05$ & $20.0$ & $0.130$ & $0.084$ & $0.047$ \\
$0.10$ & $10.0$ & $0.184$ & $0.128$ & $\mathbf{0.057}$ \\
$0.12$ & $ 8.3$ & $0.202$ & $0.142$ & $0.060$ \\
$0.15$ & $ 6.7$ & $0.226$ & $0.163$ & $0.063$ \\
$0.20$ & $ 5.0$ & $0.261$ & $0.193$ & $0.067$ \\
$0.25$ & $ 4.0$ & $0.292$ & $0.221$ & $0.071$ \\
$0.30$ & $ 3.3$ & $0.319$ & $0.246$ & $0.074$ \\
\hline
\end{tabular}
\end{table}

\noindent
The window $C_{\rm crit}^{(3B)} < C < C_{\rm crit}^{(2B)}$ exists for
all $m_{\rm sp} \lesssim 0.4\,l_{\rm Pl}^{-1}$ (range $\lambda\gtrsim 2.5\,l_{\rm Pl}$).
Its width \emph{grows} monotonically from $\Delta C \approx 0.047$ at
$m_{\rm sp}=0.05$ to $\Delta C \approx 0.074$ at $m_{\rm sp}=0.30$
(see Table~\ref{tab:Ccrit_window}).

\textbf{Numerical note:} Values for $m_{\rm sp}\geq0.15$ were computed
using a dense two-step scan over $d\in[0.3,\,25]\,l_{\rm Pl}$ (600
coarse + 500 refinement points) rather than scalar optimization, which
fails to locate the shallow minimum at $d\sim4$--$8\,l_{\rm Pl}$ for
short-range Yukawa potentials.

\subsection{The 3B-induced window at $m_{\rm sp}=0.1$}

For $m_{\rm sp}=0.1\,l_{\rm Pl}^{-1}$ ($\lambda=10\,l_{\rm Pl}$):
\begin{equation}
  C_{\rm crit}^{(3B)} = 0.128, \qquad
  C_{\rm crit}^{(2B)} = 0.184.
\end{equation}
In the interval $C\in(0.128,\,0.184)$:

\begin{table}[h]
\centering
\caption{3B-induced bound states for $m_{\rm sp}=0.1$, $C\in(C_{\rm crit}^{(3B)},\,C_{\rm crit}^{(2B)})$.}
\label{tab:3b_induced}
\begin{tabular}{rcccc}
\hline
$C$ & $E_{\rm min}^{(2B)}$ & $E_{\rm min}^{(2B+3B)}$ & $d_{\rm eq}$ [$l_{\rm Pl}$] & $E_{\rm bind}$ [$E_{\rm Pl}$] \\
\hline
$0.141$ & $+0.00032$ & $-0.00248$ & $5.74$ & $0.0025$ \\
$0.155$ & $+0.00030$ & $-0.00714$ & $4.50$ & $0.0071$ \\
$0.169$ & $+0.00027$ & $-0.01452$ & $3.70$ & $0.0145$ \\
$0.183$ & $+0.00006$ & $-0.02525$ & $3.13$ & $0.0253$ \\
\hline
\end{tabular}
\end{table}

\noindent
The equilibrium separation $d_{\rm eq}\approx3$--$6\,l_{\rm Pl}$ places
the triangle deep in the sub-Planck regime.

\subsection{Connection to Efimov physics}

This phenomenon is the TGP analog of \textbf{Efimov states}
\cite{Efimov1970}: in nuclear/atomic physics, three identical bosons
can form an infinite tower of bound states even when their pairwise
interaction is too weak to support a two-body bound state.  The
mechanism is a $1/R^2$ effective potential in the three-body channel
(where $R$ is the hyper-radius), which overcomes the centrifugal barrier.

In TGP, the analogous mechanism is:
\begin{itemize}
\item Two-body Yukawa $V_2 = -C^2 e^{-m_{\rm sp}d}/d$ cannot overcome
  quantum pressure for $C < C_{\rm crit}^{(2B)}$.
\item The Feynman three-body integral $V_3 = -C^3 I_Y(d,d,d;\,m_{\rm sp})$
  provides additional attractive energy scaling as $C^3$ (growing faster
  than $V_2\sim C^2$ as $C$ increases) which can \emph{tip the balance}.
\item Result: an equilateral triangle is bound at $d_{\rm eq}\sim3$--$6\,l_{\rm Pl}$
  in a range of $C$ where no two-body bound state exists.
\end{itemize}

\subsection{Physical parameters of the 3B-induced state}

At $m_{\rm sp}=0.1\,l_{\rm Pl}^{-1}$, $C=0.155$ (midpoint of window):
\begin{align}
  d_{\rm eq} &\approx 4.5\,l_{\rm Pl} \approx 7\times10^{-35}\,\text{m}, \\
  E_{\rm bind} &\approx 7\times10^{-3}\,E_{\rm Pl} \approx 1.4\times10^7\,\text{J}, \\
  m_{\rm body} &= 2\sqrt{\pi}\,C\,m_{\rm Pl} \approx 1.2\times10^{-8}\,\text{kg}
    \approx 0.55\,m_{\rm Pl}.
\end{align}

The 3B-induced window in mass space (for $m_{\rm sp}=0.1$):
\begin{equation}
  m_{\rm body} \in [9.8\times10^{-9},\,1.4\times10^{-8}]\,\text{kg}
  \quad(= [0.45,\,0.65]\,m_{\rm Pl}).
\end{equation}

\subsection{WKB level count: finite tower (ex24)}

Using the WKB quantisation condition
$n(E) = \frac{1}{\pi}\int_{d_L}^{d_R}\sqrt{2\mu(E-V_{\rm eff})}\,dd = n+\tfrac{1}{2}$,
with $V_{\rm eff}(d)=\frac{3}{8d^2}+V_2+V_3$ and
$\mu=1\,m_{\rm Pl}$:

\begin{table}[h]
\centering
\caption{WKB bound-state count in the 3B window, $m_{\rm sp}=0.1$.}
\label{tab:wkb_count}
\begin{tabular}{rcccr}
\hline
$C$ & $N_{\rm states}$ & $E_0$ [$E_{\rm Pl}$] & $d_{\rm eq}$ [$l_{\rm Pl}$] & \\
\hline
$0.128$--$0.175$ & 0 & --- & 5--8 & (below WKB threshold) \\
$0.175$ & 1 & $-0.00029$ & $3.48$ & \\
$0.180$ & 1 & $-0.00090$ & $3.26$ & \\
$0.184$ & 1 & $-0.00156$ & $3.04$ & \\
\hline
\end{tabular}
\end{table}

For $C=0.180$, the single WKB ground state has:
\begin{equation}
  E_0 = -9\times10^{-4}\,E_{\rm Pl},\qquad
  d_L = 1.9\,l_{\rm Pl},\qquad
  d_R = 16.3\,l_{\rm Pl},\qquad
  \Delta d = 14.4\,l_{\rm Pl}.
\end{equation}
The state is very \emph{diffuse} (spread over $\sim14\,l_{\rm Pl}$) and weakly bound.

\textbf{Key finding:} TGP supports \emph{at most one} WKB-level three-body bound
state in the window $C\in(C_{\rm crit}^{(3B)},\,C_{\rm crit}^{(2B)})$.
There is no infinite Efimov tower.

\textbf{Why no tower:} The classical Efimov effect arises from an exact $\sim1/R^2$
hyper-radial potential, which supports an infinite geometric series of levels.
The TGP three-body integral $V_3(d)\sim -C^3 I_Y(d)$ is \emph{screened}
(decays as Yukawa for $d\gtrsim\lambda=1/m_{\rm sp}$), producing only a
finite potential well and at most one or two bound levels.

\medskip
\noindent\textbf{Erratum (ex26, Sec.~\ref{sec:quantum_window}):}
The WKB result above is \emph{invalid} for this regime.
At $C=0.180$ the de~Broglie wavelength at the well centre is
$\lambda_{\rm dB}\approx44\,l_{\rm Pl}\gg L_{\rm well}\approx14\,l_{\rm Pl}$,
violating the WKB validity condition.
Exact finite-difference diagonalisation (ex26) shows the state
is \emph{unbound} ($E_0^{\rm FD}=+1.8\times10^{-3}\,E_{\rm Pl}>0$)
for all $C\leq0.184$.  The true quantum Efimov window is
$C\in(0.191,\,0.310)$; see Sec.~\ref{sec:quantum_window}.

\subsection{Falsifiability}

This is a \textbf{falsifiable prediction} of TGP:

\begin{enumerate}
\item If $m_{\rm sp}\in(0.05,\,0.4)\,l_{\rm Pl}^{-1}$ (scalar range
  $\lambda\in(2.5,20)\,l_{\rm Pl}$), triangular bound states of
  Planck-mass objects should exist that are absent in the two-body theory.
\item The window width $\Delta C\approx0.05$ translates to a mass window
  $\Delta m/m_{\rm Pl}\approx0.35$ — not fine-tuned.
\item No such state should exist for ordinary matter
  ($C\sim10^{-20}\ll C_{\rm crit}^{(3B)}$).
\end{enumerate}

\begin{table}[h]
\centering
\caption{Summary: TGP Efimov-analog bound states.}
\label{tab:efimov_summary}
\begin{tabular}{ll}
\hline
Property & Value \\
\hline
Window existence & $m_{\rm sp} < 0.4\,l_{\rm Pl}^{-1}$, $\lambda > 2.5\,l_{\rm Pl}$ \\
Window width & $\Delta C \approx 0.05$ (independent of $m_{\rm sp}$) \\
$d_{\rm eq}$ (centre of window) & $3$--$6\,l_{\rm Pl}$ \\
$E_{\rm bind}$ (centre of window) & $10^{-3}$--$10^{-2}\,E_{\rm Pl}$ \\
Mass range (window, $m_{\rm sp}=0.1$) & $[9.8\times10^{-9},\,1.4\times10^{-8}]$ kg \\
Analogy & Efimov states (nuclear/atomic physics) \\
\hline
\end{tabular}
\end{table}
